---
title: "p8105_hw1_xz2788"
author: "Xiaoyue Zhang"
date: "9/15/2018"
output: github_document
---


```{r loading_package}
library(tidyverse)
```


# Section 1

This section displays the answer to problem 1

Here is a **code chunk** that creats a dataframe

```{r creating_dataframe}
hw1_df_1 = tibble(
  numeric_vec = runif(10, 0, 5),
  logic_vec = numeric_vec > 2,
  char_vec = character(length = 10),
  factor_vec = factor(c(1, 2, 2, 2, 2, 1, 1, 3, 1, 2), labels = c("I","II","III"))
)
```

take the mean of each variable in the dataframe

```{r taking_mean}
mean(hw1_df_1$numeric_vec)
mean(hw1_df_1$logic_vec)
mean(hw1_df_1$char_vec)
mean(hw1_df_1$factor_vec)
```

Results of taking the mean of each variable:

We can take the mean of a random sample from uniform distribution and a logical vector, but it didn't work with a character vector and a factor vector.

Explanation:

In R, mean of a vector is the average or arithmetic mean of elements in the vector.

A random sample of a uniform distribution is a numerical vector. So you can take the mean of this vector. 

For a logical vector, TRUE is considered as 1 and FALSE is considered as 0 in R. So you can still take the arithmetic mean of these elements.

For a character vector, each element in the vector is a string of one or more characters instead of being considered as numbers. So R can't take the mean of characters.

Factors in R are stored as a vector of integer values with a corresponding set of character values to use when the factor is displayed. Factor's levels are always in character values and R can't take the mean of characters.

Here is a **code chunk** that applies the as.numeric function:

```{r apply_as.numeric, eval=FALSE}
## why can't use results = hide, the message said that hide not found
as.numeric(hw1_df_1$logic_vec)
as.numeric(hw1_df_1$char_vec)
as.numeric(hw1_df_1$factor_vec)
```

What happens?

The character vector was presented with NAs. The logical vector displays "1" for TRUE and "0" for FALSE. The factor vector was presented with corresponding value of levels.

Here is a **code chunk** that converts character variable to factor to numeric:

```{r convert_char_to_numeric}
char_to_factor = as.factor(hw1_df_1$char_vec)
as.numeric(char_to_factor)
```

Here is a **code chunk** that converts factor to character to numeric:

```{r convert_factor_to_numeric}
factor_to_char = as.character(hw1_df_1$factor_vec)
as.numeric(factor_to_char)
```

What happens through converting?

All the elements in character vector is converted to numerical variables with value of "1" and the factor vector is converted to a numerical vector with NAs. Because the character is considered as only a character instead of numbers in R and the factors can convert to numerical according to its corresponding levels.


# Section 2

This section displays the answer to problem 2

```{r creating_dataframe_2}
hw1_df_2 = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  logic_vec = x + y > 0,
  numeric_vec = as.numeric(logic_vec),
  factor_vec = as.factor(logic_vec)
)
```

The size of the dataset is `r dim(hw1_df_2)` or `r nrow(hw1_df_2)` * `r ncol(hw1_df_2)`

The mean of x is `r mean(hw1_df_2$x)` and the median of x is `r median(hw1_df_2$x)`

The proportion of cases for which the logical vector is TRUE is `r mean(hw1_df_2$logic_vec)`

Making scatterplots:

```{r scatterplots}
ggplot(hw1_df_2, aes(x = x, y = y, color = logic_vec)) + geom_point()
ggsave("scatter_plots_1.pdf", height = 4, width = 6)
ggplot(hw1_df_2, aes(x = x, y = y, color = numeric_vec)) + geom_point()
ggplot(hw1_df_2, aes(x = x, y = y, color = factor_vec)) + geom_point()
```
